---
title: "Regression Models Project - Motor Trend MPG Data Analysis"
author: "Banu Priya"
date: "Aug 18, 2019"
output:
  pdf_document: default
  html_document: default
---

##-- Executive Summary -->
This report showcases the analysis of the relationship between transmission type (manual or automatic) and miles per gallon (MPG). The report is set out to determine which transmission type produces a higher MPG. The `mtcars` dataset is used for this
analysis. A t-test between automatic and manual transmission vehicles shows that the
manual transmission vehicles have a 7.245 greater MPG than the automatic 
transmission vehicles. On fitting the multiple linear regressions, the analysis 
shows that the manual transmission contributed less significant to MPG & also
an improvement of 1.81 MPG.

##-- Load Data -->
Load the dataset and convert categorical variables to factors.
```{r results='hide', message=FALSE}
library(ggplot2)
data(mtcars)
head(mtcars, n=3)
dim(mtcars)
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$vs <- as.factor(mtcars$vs)
mtcars$am <- factor(mtcars$am)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
attach(mtcars)
```

##-- Exploratory Analysis -->
Exploratory Box graph that compares Automatic and Manual 
transmission MPG. The graph leads us to believe that there is a significant 
increase in MPG when for vehicles with a manual transmission vs automatic.

```{r echo=FALSE}
  boxplot(mpg ~ am, 
          xlab="Transmission Type (0 = Automatic, 1 = Manual)", 
          ylab="MPG",
          main="MPG by Transmission Type")
```

##-- Statistical Inference -->
T-Test transmission type and MPG
```{r}
testResults <- t.test(mpg ~ am)
testResults$p.value
```
The T-Test rejects the null hypothesis that the difference between transmission
types is 0.  
```{r}
testResults$estimate
```
The difference estimate between the 2 transmissions is 7.24494 MPG in favor of 
manual.

##-- Regression Analysis -->
Fit the full model of the data
```{r results='hide'}
fullModelFit <- lm(mpg ~ ., data = mtcars)
summary(fullModelFit)  # results hidden
summary(fullModelFit)$coeff  # results hidden
```
Since none of the coefficients have a p-value less than 0.05 we cannot conclude
which variables are more statistically significant. 

Backward selection to determine which variables are most statistically 
significant
```{r results='hide'}
stepFit <- step(fullModelFit)
summary(stepFit) # results hidden
summary(stepFit)$coeff # results hidden
```

The new model includes 4 variables (cylinders, horsepower, weight, transmission). The
R-squared value of 0.8659 confirms that this model explains about 87% of the 
variance in MPG. The p-values also are statistically significant because they
have a p-value less than 0.05. The coefficients conclude that increasing the 
number of cylinders from 4 to 6 with decrease the MPG by 3.03.  Further 
increasing the cylinders to 8 with decrease the MPG by 2.16. 

##-- Residuals & Diagnostics -->
Residual Plot
```{r echo=FALSE}
par(mfrcol = c(2, 2))
plot(stepFit)
```

The plots conclude:

a) The randomness of the Residuals vs. Fitted plot supports the assumption of
independence
b) The points of the Normal Q-Q plot following closely to the line conclude that
the distribution of residuals is normal
c) The Scale-Location plot random distribution confirms the constant variance 
assumption
d) Since all points are within the 0.05 lines, the Residuals vs. Leverage 
concludes that there are no outliers
```{r}
sum((abs(dfbetas(stepFit)))>1)
```

##-- Conclusion -->
There is a difference in MPG based on the transmission type. A manual transmission
will have a slight MPG boost. However, it seems that weight, horsepower, & 
number of cylinders are more statistically significant while determining MPG.
